Testing and Isolation Efficacy; Insights from a Simple Epidemic Model

### 2021-02-03 SIR_manuscript.tex
works on discussion and supp material

\begin{equation}
\label{R0}
\Rnum= (A \times S_u^* + B \times S_n^*) \times C, 
\end{equation}
where
\begin{align*}
A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
\end{align*}
Further details of derivation of $\Rnum$ are provided in the Supplementary material section.

The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref?) by using simple linear Algebra package. 
  We used R for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in R. The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing. The parameter values or ranges are as follows (I like using a table here, with the param, their units and values). It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}. This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal. For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy. Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; \bmb{remind the reader about the parameters and their values}. The solid line in each panel represents the critical contour of $\Rnum=1$. \bmb{we still have some collisions in the y-axis tick labels}}
\label{pan}
\end{figure}

@@

The model is 
\begin{subequations}\label{model}
\begin{align}
 d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
 d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
 d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
 d S_c/dt &= -\Lambda S_c + \omega S_p, \\
 d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u, \\
 d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
 d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
 d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c, \\
 d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
 d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n, \\
 d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p, \\
 d R_c/dt&= \gamma I_c + \omega R_p, \\
 dN/dt &= \omega (S_n + I_n + R_n),  \\
 dP/dt &= \omega(I_p + R_p) ,
\end{align}
\end{subequations}

where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_c$ are the isolation parameters for awaiting and reported individuals, respectively. $\Lambda$ is the force of infection defined in Eq. \eqref{Lambda}.

The next generation matrix is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible. For model, \crefrange{eq1}{eq12}, $F$ and $V$ are

@@

The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
\begin{equation}
G = \left[ \begin {array}{cc}
G_{11}&G_{12}\\
0&0
\end {array} \right], \text{ where } \\
G_{11} =C
\left[\begin {array}{cc}
A\,S_u & B\,S_u\\
A\,S_n & B\,S_n
\end {array}\right].
\end{equation}

The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.

>>>>

\begin{equation}
\label{R0}
\Rnum= (A \times S_u^* + B \times S_n^*) \times C, 
\end{equation}
where
\begin{equation}
\begin{aligned}
\label{eq:abc}
A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
\end{aligned}
\end{equation}
Further details of derivation of $\Rnum$ are provided in the Supplementary material section.

The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package (ref?) in R. The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing. The parameter values or ranges are as follows (I like using a table here, with the param, their units and values). It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}. This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal. For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy. Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at differnt levels of testing and isolation.}
We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):
$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta_w \in [0,1]$, $\eta_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $W_S=W_I=W_R=1$ and (b) targeted testing where $W_S=0.3$ and $W_I=W_R=1$. The black solid line in each panel represents the critical contour of $\Rnum=1$. \bmb{we still have some collisions in the y-axis tick labels}}
\label{pan}
\end{figure}

@@

The model in the form of a system of ordinary differential equations is 
\begin{subequations}\label{model}
\begin{align}
 d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
 d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
 d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
 d S_c/dt &= -\Lambda S_c + \omega S_p, \\
 d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u, \\
 d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
 d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
 d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c, \\
 d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
 d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n, \\
 d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p, \\
 d R_c/dt&= \gamma I_c + \omega R_p, \\
 dN/dt &= \omega (S_n + I_n + R_n),  \\
 dP/dt &= \omega(I_p + R_p) ,
\end{align}
\end{subequations}

where parameters are specified in Table \ref{tab:params}. The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible. 
Matrices $F$ and $V$ are

@@

The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
\begin{equation}
\label{mat:G}
G = \left[ \begin {array}{cc}
G_{11}&G_{12}\\
0&0
\end {array} \right], \text{ where } \\
G_{11} =C
\left[\begin {array}{cc}
A\,S_u & B\,S_u\\
A\,S_n & B\,S_n
\end {array}\right],
\end{equation}
where expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
Also note that, all $S_u$'s and $S_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
